---
title: investigate the exponential distribution in R and compare it with the Central
  Limit Theorem
author: "ismael"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

```{r, overview, echo=FALSE}
# load required libraries
library(knitr)
library(ggplot2)
library(stringr)
library(tidyverse)
library(rmarkdown)
```

## Simulations

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials

```{r, simulations, echo=FALSE}
set.seed(1)
lambda = 0.2
exponentials = 40
```

Run Simulations with variables
```{r, echo=FALSE}
simMeans = NULL
for (i in 1 : 1000) simMeans = c(simMeans, mean(rexp(exponentials, lambda)))
```

#### Calculate Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r, echo=FALSE}
mean(simMeans)
```

#### Calculate Theoretical Mean
Calculating the theoretical mean of an exponential distribution (lambda^-1).
```{r, echo=FALSE}
lambda^-1
```

#### Comparison of means
Sample Mean and Theoretical Mean are almost the same
```{r, echo=FALSE}
abs(mean(simMeans)-lambda^-1)
```

#### Calculate Sample Variance
Calculating the variance from the simulation means with give the sample
variance.
```{r, echo=FALSE}
var(simMeans)
```

#### Calculate Theoretical Variance
The theoretical variance of an exponential distribution is 
(lambda * sqrt(n))^-2.
```{r, echo=FALSE}
(lambda * sqrt(exponentials))^-2
```

#### Comparison of Variances
Sample Variance and Theoretical Variance are almost the same
```{r, echo=FALSE}
abs(var(simMeans)-(lambda * sqrt(exponentials))^-2)
```

## Distribution
```{r, message=FALSE, warning=FALSE}
simnum <- 1000
sim <- matrix(rexp(simnum*exponentials, rate=lambda), simnum, exponentials)
row_means <- rowMeans(sim)
# plot of histogram of means
hist(row_means, breaks=100, prob=TRUE,
     main="Normal Distribution Comparison", xlab="Sample mean")
# density of the means of samples
lines(density(row_means), col="blue")
# theoretical center of distribution
abline(v=1/lambda, col="red")
# theoretical density of the means of samples
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(exponentials)))
lines(xfit, yfit, pch=22, col="green", lty=2)
# add legend
legend('bottomright', c("simulation", "theoretical"), lty=c(1,2), col=c("blue", "green"))
```

Due to the Central Limit Theorem, the distribution of means of the sampled exponential distributions follow a normal distribution.